\sysfull

### 2018-03-22 paper/background.tex
sec 2
some TODOs for oopsla18

\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}. We assume a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM. Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}. \sys supports I/O behaviour similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.

@@

@@

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}                             & Reg. + Stack     \\
		DINO~\cite{dino}                                     & Reg. + WAR variables \\
		Chain~\cite{chain}                                   & PC   + Channel data\\
		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
		\hline
	\end{tabular}
	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

@@

\textbf{Memory Consistency Enforcement Overhead.} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf. \cite[Sec. 2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.

@@

Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity. Buffer sizes from prior work vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer. The key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer. \sys's {\em task coalescing}, described
in Section~\ref{sec:task_coalescing}, is motivated by these observations.

>>>>

\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}. We assume a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM. Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. We do not require a particular non-volatile memory technology, nor do we require architectural additions to commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}. \sys supports I/O behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.

@@

\xxx{reference work by Intel}

@@

\begin{table}
	\centering
	\footnotesize
	\begin{tabular}{|c|c|}
		\hline
		Model & Data Copied to/from NVRAM \\
		\hline\hline
		Mementos~\cite{mementos}                             & Reg. + Stack     \\
		DINO~\cite{dino}                                     & Reg. + WAR variables \\
		Chain~\cite{chain}                                   & PC   + Channel data\\
		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
        \cite{intel} & TODO \\ 
		\hline
	\end{tabular}
	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
	\label{table:chechpoint_comparison}
\end{table}

@@

\textbf{Memory Consistency Enforcement Overhead.} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. \xxx{Intel: undo and redo logging cost} \xxx{todo for table}. Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf. \cite[Sec. 2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.

@@

Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity. Buffer sizes from prior work vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer. The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer
either risks non-termination or poor performance due to high overhead. \sys's {\em task coalescing}, described
in Section~\ref{sec:task_coalescing}, is motivated by these observations.

### 2018-03-22 paper/intro.tex
some TODOs for oopsla18

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.

@@

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage. Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size. However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution. Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.

@@

Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate effort.

@@

\begin{itemize}
	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
	\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coalescing at runtime.
\end{itemize}

>>>>

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl} \xxx{tictpl is not HW, is it? also it doesn't deal with inconsistency; this paragraph doesn't include Hibernus, so it should not include CPTL either.} New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.

@@

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
\xxx{add \cite{intel} work; separate by-programmer and by-compiler; both cases
hit the energy estimation problem; also intel work proposes a method for decomposing
into one of many decompositions, but does not give a good solution for choosing one}
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage. Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size~\cite{cleancut}.
However, statically predicting energy of arbitrary input-dependent code with
peripheral access is a problem without a general solution. Furthermore, a
static decomposition approach prevents portability across devices with
different storage capacitors.

@@

Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. \xxx{Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate effort.}

@@

\begin{itemize}
	\item \xxx{Two} adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
	\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coalescing at runtime.
\end{itemize}

### 2018-03-22 paper/systemdescription.tex
some TODOs for oopsla18

\sys is a new programming and execution model for intermittent computing on energy-harvesting devices. \sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em accessible}, {\em efficient} and {\em flexible}. \sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.

>>>>

\sys is a new programming and execution model for intermittent computing on
energy-harvesting devices. \sys addresses the challenges outlined in
Section~\ref{sec:background} to make task-based intermittent programs {\em
accessible}, {\em efficient} and {\em flexible}. \sys accomplishes this goal
with a new program analysis and run time system for intermittent applications,
refer to Figure~\ref{fig:system_overview}.
